---
title: "Example"
author: "JL"
date: "7/23/2018"
output: 
  html_document: 
    keep_md: yes
---

## Example workflow

1. Select "Tsne Plots"
1. Color middle tsne plot by "naive_ClusterFreq" using the "Middle Plot Color" dropdown to see which meta clusters are naive cells
2. Meta clusters 1, 2, 8, 9, and 16 seem to be made up of mostly naive cells (less so 8)
3. Select meta clusters 1, 2, 8, 9, and 16 in the "Meta clusters to characterize" dropdown
4. Select "Raw Marker Heatmap" to see which markers differentiate these meta-clusters
  - Seems to be IgD, CD95, CD19, maybe CD4
5. Similar process for "Normalized Heatmap"
6. Select "Marker Distributions"
  - Select IgD, CD95, CD19, and CD4 in the "Markers" dropdown
  - Resonably separated distributions
8. Select "Population Freq Heatmap", can confirm that cluster 8 is a little less naive-like
7. Go back to "Tsne Plots" and color a plot by IgD 

### Description of tabs



```{r echo=FALSE}

d=read.delim("../report/tabDesc.txt",stringsAsFactors = FALSE)
library(knitr)
kable(d,format = "markdown")

```


### TODO

1. legends!! Currently the legends are placed over the plots of interest
1. add p-values to distribution plots

### Current Methods


1. Run phenograph on each individual sample, starting with cytotoxic T cells
  - Results displayed here are from CTL samples run on LSR only
2. For each sample's phenograph cluster compute:
  1.  median compensated/transformed expression across all markers used (calling this "RAW")
  2.  median of the normalized (mean 0, SD 1) compensated/transformed expression across all markers use
  3. Frequency of known cytotoxic subsets in the cluster
  4.  Proportion of events in the phenograph cluster out of all events in each known cytotoxic subset
3. Combine individual sample's centroids into a single matrix
4. Run phenograph on the combined matrix (currently using "RAW")
5. Run tsne on the combined matrix (currently using "RAW")
6. Plot results, coloring by things

### Data column descriptions

- just for reference

```{r echo=FALSE}

d=read.delim("../report/summary.columns")
library(knitr)
kable(d,format = "markdown")

```

