---
title: "Insilco LOD"
author: "JL"
date: "01/08/2018"
output: 
  html_document: 
    keep_md: yes
---


# Notes

Selected 30 LSR and 30 FORTESSA files matched for panel1 and panel2 that:

- had an event count > 700000
- had 0 "bad","abnormal", or "manual" annotatations

Within this set:

- randomly sub-sampled events 10 times per threshold, with requirement that number of events sampled be < %50 of total events





```{r setup, include=FALSE,eval=FALSE}
totals = read.delim("/Volumes/Beta/data/flow/LODCtrlsV10/all.totalCellCounts.metrics.txt", stringsAsFactors = FALSE)
freqs = read.delim("/Volumes/Beta/data/flow/LODCtrlsV10/all.freq.metrics.txt", stringsAsFactors = FALSE)
totals$SUCCESS=totals$FILE %in% freqs$name
totals$GROUP=gsub("_iter.*","",totals$FILE)
lymph=freqs[which(freqs$Manual=="Lymphocytes (SSC-A v FSC-A)"),]
lymph$LYMPH_COUNT=lymph$Count
lymph=lymph[,c("name","LYMPH_COUNT")]
pbmc=freqs[which(freqs$Manual=="Live Single PBMCs (SSC-A/FSC-A)"),]
pbmc$PBMC_COUNT=pbmc$Count
pbmc=pbmc[,c("name","PBMC_COUNT")]

merge=merge(totals,freqs,by.x ="FILE",by.y = "name",all.x=TRUE )
merge=merge(merge,lymph,by.x ="FILE",by.y = "name",all.x=TRUE )
merge=merge(merge,pbmc,by.x ="FILE",by.y = "name",all.x=TRUE )


save(merge,file ="/Volumes/Beta/data/flow/LODCtrlsV10/merge.RData" )
```

```{r  include=FALSE}

load("/Volumes/Beta/data/flow/LODCtrlsV10/merge.RData")
#24HRS
#over
#_BC-D24-S+F_72HRS
#Ctl
dontCare <- c("24HRS", "over","D24","SEPM","DAYS","ASMIC","bad","P1","baseline")
nonHZ=!grepl(paste(dontCare,collapse="|"),merge$FILE)

merge=merge[nonHZ,]
mergeAll=merge
# use=grepl("Ctl",merge$FILE)
# merge=merge[use,]
# use=grepl("iter_1",merge$FILE)
# merge=merge[use,]


# merge[which(merge$SUCCESS==FALSE&merge$TOTAL_COUNTS>100000),]$FILE

# use=grep("_F",merge$FILE)
# merge=merge[use,]
fail = merge[which(!merge$SUCCESS),]

merge$ID=gsub(".*_F","F",merge$FILE)
merge$ID=gsub("_.*","",merge$ID)

merge$ORIGINAL_COUNT = "FALSE"
original = grep("All",merge$FILE)
merge$ORIGINAL_COUNT[original]=TRUE

CV <- function(mean, sd){
      (sd/mean)*100
}

cvCalc <- function(vec){
  CV(mean = mean(vec,na.rm = TRUE),sd=sd(vec,na.rm = TRUE))
}

library(irr)
krpAlpha <- function(vec){
  # print(class(vec))
  df=as.matrix(vec)
  # print(length(vec))
  if(length(vec)<=1){
    return(NA)
  }
  t=kripp.alpha(df ,"ratio")
  return(t$value)
}
mergeAll$CV_POP_KEY=paste0(mergeAll$GROUP,"_",mergeAll$Manual)
cvResults=aggregate(mergeAll$freqParent, list(mergeAll$CV_POP_KEY), cvCalc)
colnames(cvResults)=c("CV_POP_KEY","SAMPLE_CV")
mergeAll=merge(mergeAll,cvResults,by.x="CV_POP_KEY",by.y="CV_POP_KEY")

# kaResults=aggregate(mergeAll$freqParent, list(mergeAll$CV_POP_KEY), krpAlpha)
# colnames(kaResults)=c("CV_POP_KEY","SAMPLE_KA")
# mergeAll=merge(mergeAll,kaResults,by.x="CV_POP_KEY",by.y="CV_POP_KEY")

mergeAll$ORIGINAL_COUNT = FALSE
original = grep("All",mergeAll$FILE)
mergeAll$ORIGINAL_COUNT[original]=TRUE
mergeAll=mergeAll[which(mergeAll$ORIGINAL_COUNT=="FALSE"),]

# merge[merge$ORIGINAL_COUNT,]$TOTAL_COUNTS
# merge=merge[which(merge$COUNTS<=merge$TOTAL_COUNTS),]

library(ggplot2)
t2 <- theme(
  axis.line = element_line(colour = "black"),
  axis.text = element_text(colour = "black"),
  axis.ticks = element_line(colour = "black"),
  # panel.grid.major.x = element_blank(),
  panel.grid.minor.x = element_blank(),
  # panel.grid.major.y = element_blank(),
  # panel.grid.minor.y = element_blank(),
  panel.border = element_blank(),
  # panel.background = element_blank(),
  # legend.position="none",
  axis.text.x=element_text(angle=90,hjust=1)
)
theme_set(theme_grey(base_size = 18)) 


```

# Failure rate by number of events sampled
```{r fig.height=10,fig.width=12,echo=FALSE}

failData=merge[which(!merge$SUCCESS|merge$Manual=="Tcells (CD3+ CD19-)"|merge$Manual=="MONOCYTES (CD14+)"),]
failData=failData[which(failData$ORIGINAL_COUNT=="FALSE"),]

plotFailData = data.frame(EVENT_COUNT=numeric(),FAIL=character(),NUMBER=numeric(),PANEL=character())
for(panel in unique(failData$PANEL) ){
  if(panel!="panel_undetermined"){
  for(machine in unique(failData$MACHINE.x)){
  tmp = failData[which(failData$PANEL==panel&failData$MACHINE.x==machine),]
  t=as.data.frame(table(tmp$TOTAL_COUNTS,is.na(tmp$Manual)))
  colnames(t)=c("EVENT_COUNT","FAIL","NUMBER");
  t$PANEL=panel

  pbf=ggplot(t,aes(x=EVENT_COUNT,y=NUMBER,colour=FAIL,size=FAIL)) +
  geom_point() +
  xlab("Num Events Sampled") + 
  ylab(paste0("COUNT"))+t2+ggtitle(paste0(panel, " ", machine)) 
  # print(pbf)
  
  rate=data.frame(EVENT_COUNT=unique(t$EVENT_COUNT))
  failCounts = t[which(t$FAIL==TRUE),]
  failCounts$NUM_FAILS=failCounts$NUMBER
  successCounts= t[which(t$FAIL==FALSE),]
  successCounts$NUM_SUCCESS=successCounts$NUMBER
  
  rate=merge(rate,failCounts,by.x="EVENT_COUNT",by.y="EVENT_COUNT",all.x=TRUE)
  rate=merge(rate,successCounts,by.x="EVENT_COUNT",by.y="EVENT_COUNT",all.x=TRUE)
  rate$FAIL_RATE=rate$NUM_FAILS/(rate$NUM_FAILS+rate$NUM_SUCCESS)
  rate$FAIL_RATE =rate$FAIL_RATE*100
  pbfr=ggplot(rate,aes(x=EVENT_COUNT,y=FAIL_RATE)) +
  geom_point() +
  xlab("Num Events Sampled") + 
  ylab(paste0("Percent failure"))+t2+ggtitle(paste0(panel, " ", machine)) +ylim(0,100)
  print(pbfr)
  }
  }
}

```

# Population specific variability


- Frequency Parent CV computed within a sample for each event threshold (10 replicates at 2k,10 replicates at 3k, etc)

Plots below - for each population

1. Boxplots of sample CVs at each event threshold (the total events in the .fcs file)
2. Number of events in lymph/pbmc populations vs sample CV
4. Boxplots of sample CVs binned by lymph/pbmc counts - I'm kinda finding this to be the most informative. 
5. Number of events in lymph/pbmc populations vs number of events in current population.
6. Boxplots of sample CVs binned by number of events in current population.


```{r fig.height=10,fig.width=12,echo=FALSE}


p1Looks = c(
"Lymphocytes (SSC-A v FSC-A)",
"Single Cells (FSC-H v FSC-W)",
"Live cells (PE-)",
"Tcells (CD3+ CD19-)",
"Helper Tcells-CD4+",
"cytotoxic Tcells-CD8+",
"B cells (CD3- CD19+)",
"naive Bcells (CD27- IgD+)",
"IgD- memory Bcells (CD27+)",
"IgD+ memory Bcells (CD27+)"
)

p2Looks = c(
  "Live immune cells (CD45+ PE-)",
  "Live Single immune cells(FSC-H/FSC-W)",
  "Live Single PBMCs (SSC-A/FSC-A)",
  "DC NK MONOCYTES (CD3- CD19-)",
  "MONOCYTES (CD14+)",
  "Non classical monocytes (CD16+ CD14+)",
  "Classical monocytes (CD16- CD14+)",
  "DC NK (CD20- CD14-)",
  "NK (CD16+)",
  "NK CD56LO",
  "NK CD56HI"
  )

p1full=freqs = read.delim("/Users/Kitty/git/auto-fcs/explore/openCyto/panel1Map.txt", stringsAsFactors = FALSE)$Manual[1:2]
p2full=freqs = read.delim("/Users/Kitty/git/auto-fcs/explore/openCyto/panel2Map.txt", stringsAsFactors = FALSE)$Manual[1:2]
# [1:2]
# plot(mergeAll$Count,mergeAll$SAMPLE_CV)
# plot(mergeAll$freqParent,mergeAll$SAMPLE_CV)
panels=c("panel1","panel2")
for(panel in panels){
  pops=c()
    if(panel=="panel2"){
      pops=p2full
    }
   if(panel=="panel1"){
      pops=p1full
    }

  for(popP in pops){
  plot=1  

    sub= mergeAll[which(mergeAll$PANEL==panel&mergeAll$Manual == popP),]
  pb=ggplot(sub,aes(x=as.factor(TOTAL_COUNTS),y=SAMPLE_CV,colour=MACHINE.x)) +
  geom_boxplot() +
  xlab("Num Events Sampled") +
  ylab(paste0("Frequency Parent CV"))+t2+ggtitle(paste0("Panel=",panel,"\nPop=",popP," plot # = ",plot))
  print(pb)
  plot=plot+1  

  #  pb=ggplot(sub,aes(x=as.factor(TOTAL_COUNTS),y=SAMPLE_KA,colour=MACHINE.x)) +
  # geom_boxplot() +
  # xlab("Num Events Sampled") +
  # ylab(paste0("Frequency Parent KA"))+t2+ggtitle(paste0("Panel=",panel,"\nPop=",popP))
  # print(pb)
  if(panel=="panel2"){
  pb=ggplot(sub,aes(x=PBMC_COUNT,y=SAMPLE_CV,colour=MACHINE.x)) +
  geom_point() +
  xlab("Events in Live Single PBMCs") +
  ylab(paste0("Frequency Parent CV"))+t2+ggtitle(paste0("Panel=",panel,"\nPop=",popP," plot # = ",plot))
  print(pb)
  plot=plot+1  


  # pb=ggplot(sub,aes(x=PBMC_COUNT)) +
  # geom_histogram(bins = 50) +
  # xlab("Events in Live Single PBMCs") +
  # ylab(paste0("Count"))+t2+ggtitle(paste0("Panel=",panel,"\nPop=",popP," plot # = ",plot))
  # print(pb)
  # plot=plot+1  

    sub$PbmcBin <- cut(sub$PBMC_COUNT, c(0,5000,10000,15000,seq(from=20000,to=200000,by = 10000),600000))
   sub$PopBin <- cut(sub$Count, seq(from=0,to=max(sub$Count),length.out = 50)) 
 pb=ggplot(sub,aes(x=PbmcBin,y=SAMPLE_CV,colour=MACHINE.x)) +
  geom_boxplot() +
  xlab("Events in Live Single PBMCs") +
  ylab(paste0("Frequency Parent CV"))+t2+ggtitle(paste0("Panel=",panel,"\nPop=",popP," plot # = ",plot))
  print(pb)
  plot=plot+1  

   pb=ggplot(sub,aes(x=PBMC_COUNT,y=Count,colour=MACHINE.x)) +
  geom_point() +
  xlab("Events in Live Single PBMCs") +
  ylab(paste0("Events in ",popP))+t2+ggtitle(paste0("Panel=",panel,"\nPop=",popP," plot # = ",plot))
  print(pb)
  plot=plot+1  

  pb=ggplot(sub,aes(x=PopBin,y=SAMPLE_CV,colour=MACHINE.x)) +
  geom_boxplot() +
  xlab(paste0("Events in ",popP)) +
  ylab(paste0("Frequency Parent CV"))+t2+ggtitle(paste0("Panel=",panel,"\nPop=",popP," plot # = ",plot))
  print(pb)
  plot=plot+1  

  
  }
    if(panel=="panel1"){
  pb=ggplot(sub,aes(x=LYMPH_COUNT,y=SAMPLE_CV,colour=MACHINE.x)) +
  geom_point() +
  xlab(" Events in Lymphocytes") +
  ylab(paste0("Frequency Parent CV"))+t2+ggtitle(paste0("Panel=",panel,"\nPop=",popP," plot # = ",plot))
  print(pb)
    plot=plot+1  

  # pb=ggplot(sub,aes(x=LYMPH_COUNT)) +
  # geom_histogram(bins = 50) +
  # xlab(" Events in Lymphocytes") +
  # ylab(paste0("Count"))+t2+ggtitle(paste0("Panel=",panel,"\nPop=",popP," plot # = ",plot))
  # print(pb)
  #   plot=plot+1  

  
  sub$LymphBin <- cut(sub$LYMPH_COUNT,  c(0,5000,10000,15000,seq(from=20000,to=200000,by = 10000),600000))
  sub$PopBin <- cut(sub$Count, seq(from=0,to=max(sub$Count),length.out = 50)) 

  pb=ggplot(sub,aes(x=LymphBin,y=SAMPLE_CV,colour=MACHINE.x)) +
  geom_boxplot() +
  xlab(" Events in Lymphocytes") +
  ylab(paste0("Frequency Parent CV"))+t2+ggtitle(paste0("Panel=",panel,"\nPop=",popP," plot # = ",plot))
  print(pb)
    plot=plot+1  

  pb=ggplot(sub,aes(x=LYMPH_COUNT,y=Count,colour=MACHINE.x)) +
  geom_point() +
  xlab(" Events in Lymphocytes") +
  ylab(paste0("Events in ",popP))+t2+ggtitle(paste0("Panel=",panel,"\nPop=",popP," plot # = ",plot))
  print(pb)
    plot=plot+1  

   pb=ggplot(sub,aes(x=PopBin,y=SAMPLE_CV,colour=MACHINE.x)) +
  geom_boxplot()  +
  xlab(paste0("Events in ",popP))+
  ylab(paste0("Frequency Parent CV"))+t2+ggtitle(paste0("Panel=",panel,"\nPop=",popP," plot # = ",plot))
  print(pb)
    plot=plot+1  

  
  
  }
  
  
  # pb=ggplot(sub,aes(x=Count)) +
  # geom_histogram() +
  # xlab("Events in Population") +
  # t2+ggtitle(paste0("Panel=",panel,"\nPop=",popP))
  # print(pb)
      
  }
}


#   
# 
# for(pop in pops){
#  p=ggplot(merge[which(merge$Manual==pop),],aes(x=as.factor(TOTAL_COUNTS),y=freqParent,colour=ID,size=ORIGINAL_COUNT)) +
#   geom_point() +
#   xlab("Num Events Sampled") +
#   ylab(paste0("Frequency Parent",pop))+t2+facet_wrap(~MACHINE.x) +ggtitle(paste0("POP=",pop))+ylim(0,1)
# 
#  print(p)
#   pb=ggplot(merge[which(merge$Manual==pop),],aes(x=as.factor(TOTAL_COUNTS),y=freqParent,colour=ID)) +
#   geom_boxplot() +
#   xlab("Num Sampled") +
#   ylab(paste0("Frequency Parent",pop))+t2+facet_wrap(~MACHINE.x)+ggtitle(paste0("POP=",pop))+ylim(0,1)
#   print(pb)
# }

```
